Symmetric Circuits for Rank Gates

### 2018-02-28 sections/symmetric-circuits.tex
Updates to the entire paper 28th Feb 2018

Structured functions give us a general framework for indexing the input sequence
to a Boolean function and associating with that function an appropriate symmetry
condition, or notion of equivalence. Let $F$ be a $(\tau, D)$-structured
function. We say $F$ is \emph{$G$-invariant} if it is constant on $G$-equivalent
input vectors, i.e. for all $G$-equivalent $f, g \in \{0,1\}^{X}$, $F(f) =
F(g)$. We say $F$ is \emph{isomorphism-invariant} if it is $\aut(\tau,
D)$-invariant. Note that a structured function with vocabulary $\tau$ is
isomorphism-invariant if, and only if, it decides a property of
$\tau$-structures.

@@

In each case a natural choice of $\tau$ (and $D$) indexes the input to the
structured function appropriately and produces the appropriate symmetry
condition. It is worth mentioning that the graph-invariant case is not strictly
reproduced using structured functions. This is because in the structured
function case the input string is indexed by all possible elements of the binary
relation (i.e. all ordered pairs over the universe). As such, the input string
of the structured function may encode any structure over the universe with a
single binary relation, i.e. a directed graph. In contrast, the graph-invariant
function discussed at the beginning of this subsection indexes the input string
using two element subsets of the universe, rather pairs of elements, an encoding
that naturally enforces the symmetry and irreflexivity properties the edge
relation. We should note, however, that symmetry and irreflexivity are
properties of directed-graph, and as such can be decided by an
isomorphism-invariant structured function over the graph vocabulary.

@@

For a gate $g \in C$, we let the \emph{index} of $g$, denoted by $\ind(g)$, be
$\ind(\Sigma(g))$. We define the relation $W \subseteq G^2$ by $W(h,g)$ if, and
only if, $h \in \img(L(g))$. For each $g \in G$ we let $W(\cdot, g) := \{h \in G
: W(h,g)\}$ and $W(g, \cdot) := \{h \in G : W(g,h)\}$. We call the elements of
$W(\cdot, g)$ the \emph{children} of $g$ and the elements of $W(g, \cdot)$ the
\emph{parents} of $g$. We also abbreviate $W(g, \cdot)$ by $H_g$.

@@

Let $\rho$ be a relational vocabulary, $\mathcal{A}$ a $\rho$-structure over an
$n$-element universe $U$ and $\gamma \in [n]^{\underline{U}}$. Let $\gamma
\mathcal{A}$ be the structure on universe $[n]$ formed by mapping $U$ in
accordance with $\gamma$. The evaluation of a $(\mathbb{B}, \rho)$-circuit $C$
of order $n$ computing a $q$-ary query $Q$ proceeds by recursively evaluating
the gates in the circuit. The evaluation of the gate $g$ for the bijection
$\gamma$ and input $\mathcal{A}$ is denoted by $C[\gamma \mathcal{A}](g)$, and
is given as follows
\begin{enumerate}
  \setlength\itemsep{0mm}
\item If $g$ is a constant gate then it evaluates to the bit given by
  $\Sigma(g)$,
\item if $g$ is a relational gate then $g$ evaluates to true iff $\gamma
  \mathcal{A} \models \Sigma(g)(\Lambda (g))$, and
\item if $g$ is an internal gate such that $\Sigma (g)$ let $L^{\gamma}_g:
  \ind(g) \rightarrow \{0,1\}$ be defined by $L^{\gamma}_g(x) = C[\gamma
  \mathcal{A}](L(g)(x))$, for all $x \in \ind(g)$. Then $g$ evaluates to true
  if, and only if, $\Sigma(g) (L^{\gamma}_g) = 1$.
\end{enumerate}
We say that $C$ defines the $q$-ary query $Q \subseteq U^q$ under $\gamma$ where
$\vec{a} \in Q$ if, and only if, $C[\gamma \mathcal{A}](\Omega (\gamma \vec{a}))
= 1$.

@@

\begin{lem}
  Let $C$ be a $(\mathbb{B}, \rho)$-circuit of order $n$ computing a $0$-ary
  query. The function computed by $C$ is isomorphism-invariant if, and only if,
  $C$ is an invariant circuit.
\end{lem}
\begin{proof}
  We first note that, since $C$ computes a $0$-ary query, there is exactly one
  output gate. We call this gate $g_o$. We can associate with each relational
  gate $g$ in $C$ a unique pair $(\vec{a}, R)$, where $R := \Sigma (g)$ is a
  relational symbol in $\tau$ and $\vec{a} := \Lambda_R(g)$ is an an element of
  $R^{[n]}$. Thus we can think of the function $F_C$ computed by the circuit as
  having it's input string indexed by the elements of $\ind(\rho, [n])$, and so
  think of $F_C$ as a $(\rho, [n])$-structured function.

@@

In order to proceed we will restrict ourselves to circuits with certain
constraints on how those gates computing non-symmetric functions can be wired
into the circuit. We call these circuits \emph{transparent} circuits. We show in
Sections~\ref{sec:transparency-and-unique-extensions} and~\ref{} that
transparency is sufficient to allow for the polynomial-time decidability of many
important circuit properties, while still keeping the circuit model powerful
enough for us to establish a translation from formulas to families of circuits.

Before we can formally define \emph{transparency} we need to define the
\emph{syntactic-equivalence} relation on the gates of a circuit. The intuition
is that two gates in a circuit are syntactically-equivalent if the circuits
underneath them are `hereditary similar'.

% In order to prove this result, Anderson and Dawar introduce the notion of a
% \emph{rigid} circuit, proving that symmetric circuits may be

@@

\begin{definition}
  Let $\tau$ be a many-sorted vocabulary and $D$ and $D'$ be $\tau$-sorted sets.
  Let $X$ be a set and $\sim$ be an equivalence relation defined on $X$. Let $Y,
  Z \subset X$. We say two functions $f : \ind(\tau,D) \rightarrow X$ and $g :
  \ind(\tau, D') \rightarrow Y$ are \emph{isomorphism-equivalent up to $\sim$}
  if there exists an isomorphism $\pi : \str{\tau, D} \rightarrow \str{\tau,
    D'}$ such that $f (\vec{x}_R) \sim g (\pi \vec{x}_R)$ for all $\vec{x}_R \in
  \ind(\tau, D)$ .
\end{definition}

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. We recursively define the equivalence relation
  \emph{syntactic-equivalence}, which we denote using the symbol `$\equiv$', on
  $G$ as follows. Suppose $g$ and $h$ are gates in $C$ such that $\Sigma (g) =
  \Sigma(h)$ and either both $g$ and $h$ are output gates and $\Omega^{-1}(g) =
  \Omega^{-1}(h)$ or neither are output gates. Suppose $g$ and $h$ are input
  gates, then $g \equiv h$ if, and only if, both $g$ and $h$ are constant gates
  or both are relational gates and $\Lambda(g) = \Lambda (h)$. Suppose $g$ and
  $h$ are internal gates and and suppose we have defined the
  syntactic-equivalence relation for all gates of height less than the height of
  either $g$ or $h$. Then $g \equiv h$ if, and only if, $L(g)$ and $L(h)$ are
  isomorphism equivalent up to $\equiv$.
\end{definition}

@@

We could think of any gate $g$ in the circuit as computing a function that maps
a given assignment to the input gates to the evaluation of the circuit at $g$
for the given assignment. While we would like to be able to identify gates that
compute the same function in this sense, it is not hard to show that deciding
this property is $\NP$-hard. The syntactic-equivalence relation is an
easier-to-compute refinement of this relation, instead identifying pairs of
gates if the circuits underneath them are hereditary similar.

We could also define a similar equivalence relation using the automorphisms of a
circuit. In this case we identify a pair of gates $(h, h')$ in a circuit $C$ of
order $n$ if, and only if, there exists $\sigma \in \sym_n$, $\pi, \pi' \in
\aut(C)$ both extending $\sigma$, and $g$ a gate in $C$, such that $h = \pi (g)$
and $h' = \pi' (g)$. We now show this relation is a refinement of
syntactic-equivalence.

\begin{remark}
  It's interesting to note that if $C$ has unique extensions then this relation
  collapses to equality, syntactic-equivalence does not. Note to self: should I
  mention this? Also, the following result is equivalent to checking for just
  $\pi $
\end{remark}

@@

  We now prove the result by induction on height. Suppose $g$ is a gate of
  height $0$. Then $g$ is either a relational or constant gate. In either case
  $\pi (g) = \pi' (g)$, and so $\pi(g)$ and $\pi'(g)$ are
  syntactically-equivalent.

  Suppose $g$ is an internal gate and suppose that for every gate $h$ of height
  less than $g$, $\pi (h)$ is syntactically-equivalent to $\pi'(h)$. We have
  that there exists $\lambda, \lambda' \in \aut(g)$ such that $\pi
  L(g)(\vec{a}_R) = L(\pi g) (\lambda \vec{a}_R)$ and $\pi' L(g)(\vec{a}_R) =
  L(\pi' g)(\lambda' \vec{a}_R)$, for all $\vec{a}_R \in ind(g)$. Then, from the
  inductive hypothesis, we have that $\pi L(g)(\vec{a}_R) \equiv \pi
  'L(g)(\vec{a}_R)$ and so $L(\pi g)(\lambda \vec{a}_R) \equiv L(\pi' g)
  (\lambda' \vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$. Thus we have that
  $L(\pi g) (\vec{a}_R) = L (\pi g) (\lambda \lambda^{-1}\vec{a}_R) \equiv
  L(\pi' g) (\lambda' \lambda^{-1}\vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$,
  and so $L(\pi g)$ is isomorphism equivalent to $L(\pi'g)$ up to
  syntactic-equivalence. We thus have that $\pi (g) $ and $\pi' (g)$ are
  syntactically-equivalent, and the result follows.

@@

We now define the \emph{unique labels} and \emph{transparency} conditions on
circuits.

\begin{definition}
  Let $C$ be a circuit and $g$ be a gate in $C$. We say $g$ has \emph{injective
    labels} if $L(g)$ is an injection. We say $g$ has \emph{unique labels} if
  $g$ has injective labels and no two gates in $\img(L(g))$ are
  syntactically-equivalent. We say $C$ has \emph{injective labels} (resp.
  \emph{unique labels}) if every gate in $C$ has injective labels (resp. unique
  labels). We say $C$ is \emph{transparent} if every gate $g$ in $C$ such that
  $\Sigma(g)$ is not a symmetric function has unique labels.
\end{definition}

@@

We note that the construction of $C'$ from $C$ in
Lemma~\ref{lem:redundant-gates} can be implemented so as to run in time
polynomial in the size of $C$. As such, for the remainder of this paper we
assume, unless stated otherwise, every circuit has no redundant gates. With this
assumption in mind we now show that every circuit with unique labels has unique
extensions.

@@

\begin{prop}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. If $C$ has unique labels then $C$ has unique
  extensions.
  \label{prop:unique-labels-unique-extensions}
\end{prop}
\begin{proof}
  Suppose $C$ has unique labels. Let $\pi_{\sigma}, \pi_{\sigma}' \in \aut(C)$
  be automorphisms extending $\sigma \in \sym_n$. We now prove that $\pi_\sigma
  = \pi_{\sigma}'$. Let $\pi := \pi_{\sigma}'\pi^{-1}_\sigma$. We have that
  $\pi$ is an automorphism of the circuit and that $\pi$ extends
  $\sigma^{-1}\sigma = e$, the identity permutation. It follows that $\pi$ fixes
  all input gates and, since for any output gate $g$, $\pi (g) = \pi
  (\Omega(\vec{a})) = \Omega(e \vec{a}) = \Omega(\vec{a}) = g$ for all $\vec{a}
  \in [n]^q$, $\pi$ fixes all output gates.

  We now prove the result by structural induction on the circuit starting from
  the output gates. Let $g$ be an output gate then $\pi (g) = g$ from the above
  argument. Let $g$ be an internal, non-output gate. We now show that if $\pi
  (g) = g$ then for all $h \in H_g$ we have $\pi (h) = h$. Suppose $\pi (g) = g$
  and let $h \in H_g$. Since $\pi (g) = g$ we have that $\pi H_{g} = H_g$, and
  so $\pi (h) \in H_g$. We have from
  Lemma~\ref{lem:permutation-extending-syntactic-equivalence} that $h$ is
  syntactically-equivalent to $\pi (h)$ and, since $g$ has unique labels, $h$ is
  the only gate in $H_g$ syntactically-equivalent to $h$. It follows that $\pi
  (h) = h$.

@@

We show in Proposition~\ref{prop:transparent-unique} that a transparent circuit
can be converted into an equivalent circuit with unique labels in time
polynomial in the size of the circuit. Since we are
working with $P$-uniform families of transparent circuits, this fact allows us
to assume, without a loss of generality, that transparent circuits have unique
labels (and so unique extensions).

>>>>

Structured functions give us a general framework for indexing the input sequence
to a Boolean function and associating with that function an appropriate symmetry
condition, or notion of equivalence. Let $F$ be a $(\tau, D)$-structured
function. We say $F$ is \emph{$G$-invariant} if it is constant on $G$-equivalent
input vectors, i.e.\ for all $G$-equivalent $f, g \in \{0,1\}^{X}$, $F(f) =
F(g)$. We say $F$ is \emph{isomorphism-invariant} if it is $\aut(\tau,
D)$-invariant. Note that a structured function with vocabulary $\tau$ is
isomorphism-invariant if, and only if, it decides a property of
$\tau$-structures.

@@

In each case a natural choice of $\tau$ (and $D$) indexes the input to the
structured function appropriately and produces the appropriate symmetry
condition. It is worth mentioning that the graph-invariant case is not strictly
reproduced using structured functions. This is because in the structured
function case the input string is indexed by all possible elements of the binary
relation (i.e. all ordered pairs over the universe). As such, the input string
of the structured function may encode any structure over the universe with a
single binary relation, i.e. a directed graph. In contrast, the graph-invariant
function discussed at the beginning of this subsection indexes the input string
using two element subsets of the universe, rather pairs of elements, an encoding
that naturally enforces the symmetry and irreflexivity properties the edge
relation.

% We should note, however, that symmetry and irreflexivity are
% properties of directed-graph, and as such can be decided by an
% isomorphism-invariant structured function over the graph vocabulary.

@@

For a gate $g \in C$, we let the \emph{index} of $g$, denoted by $\ind(g)$, be
$\ind(\Sigma(g))$. We define the relation $W \subseteq G^2$ by $W(h,g)$ if, and
only if, $h \in \img(L(g))$. For each $g \in G$ we let $W(\cdot, g) := \{h \in G
: W(h,g)\}$ and $W(g, \cdot) := \{h \in G : W(g,h)\}$. We call the elements of
$W(\cdot, g)$ the \emph{children} of $g$ and the elements of $W(g, \cdot)$ the
\emph{parents} of $g$. We also abbreviate $W(g, \cdot)$ by $H_g$. We write $W_T$
for the transitive closure of $W$.

@@

Let $\rho$ be a relational vocabulary, $\mathcal{A}$ a $\rho$-structure over an
$n$-element universe $U$ and $\gamma \in [n]^{\underline{U}}$. Let $\gamma
\mathcal{A}$ be the structure on universe $[n]$ formed by mapping $U$ in
accordance with $\gamma$. The evaluation of a $(\mathbb{B}, \rho)$-circuit $C$
of order $n$ computing a $q$-ary query $Q$ proceeds by recursively evaluating
the gates in the circuit. The evaluation of the gate $g$ for the bijection
$\gamma$ and input $\mathcal{A}$ is denoted by $C[\gamma \mathcal{A}](g)$, and
is given as follows
\begin{enumerate}
  \setlength\itemsep{0mm}
\item If $g$ is a constant gate then it evaluates to the bit given by
  $\Sigma(g)$,
\item if $g$ is a relational gate then $g$ evaluates to true iff $\gamma
  \mathcal{A} \models \Sigma(g)(\Lambda (g))$, and
\item if $g$ is an internal gate such that $\Sigma (g)$ let $L^{\gamma
    \mathcal{A}}(g): \ind(g) \rightarrow \{0,1\}$ be defined by
  $L^{\gamma\mathcal{A}}(g)(x) = C[\gamma \mathcal{A}](L(g)(x))$, for all $x \in
  \ind(g)$. Then $g$ evaluates to true if, and only if, $\Sigma(g)
  (L^{\gamma}_g) = 1$.
\end{enumerate}
We say that $C$ defines the $q$-ary query $Q \subseteq U^q$ under $\gamma$ where
$\vec{a} \in Q$ if, and only if, $C[\gamma \mathcal{A}](\Omega (\gamma \vec{a}))
= 1$.

@@

\begin{lem}
  Let $C$ be a $(\mathbb{B}, \rho)$-circuit of order $n$ computing a $0$-ary
  query. The function computed by $C$ is isomorphism-invariant if, and only if,
  $C$ is an invariant circuit.
\end{lem}
\begin{proof}
  We first note that, since $C$ computes a $0$-ary query, there is exactly one
  output gate. We call this gate $g_o$. We can associate with each relational
  gate $g$ in $C$ a unique pair $(\vec{a}, R)$, where $R := \Sigma (g)$ is a
  relational symbol in $\tau$ and $\vec{a} := \Lambda_R(g)$ is an an element of
  $R^{[n]}$. Thus we can think of the function $F_C$ computed by the circuit as
  having its input string indexed by the elements of $\ind(\rho, [n])$, and so
  think of $F_C$ as a $(\rho, [n])$-structured function.

@@

In order to proceed we will restrict ourselves to circuits with certain
constraints on how those gates computing non-symmetric functions can be wired
into the circuit. We call these circuits \emph{transparent} circuits. We show in
Sections~\ref{sec:transparency-and-unique-extensions} that transparency is
sufficient to allow for the polynomial-time decidability of many important
circuit properties, while still keeping the circuit model powerful enough for us
to establish a translation from formulas to families of circuits.

Before we can formally define \emph{transparency} we need to define the
\emph{syntactic-equivalence} relation on the gates of a circuit. The intuition
is that two gates in a circuit are syntactically-equivalent if the circuits
underneath them are `hereditary similar'.
% In order to prove this result, Anderson and Dawar introduce the notion of a
% \emph{rigid} circuit, proving that symmetric circuits may be

@@

% \begin{definition}
%   Let $\tau$ be a many-sorted vocabulary and $D$ and $D'$ be $\tau$-sorted
%   sets. Let $X$ be a set and $\sim$ be an equivalence relation defined on $X$.
%   Let $Y, Z \subset X$. We say two functions $f : \ind(\tau,D) \rightarrow X$
%   and $g : \ind(\tau, D') \rightarrow Y$ are \emph{isomorphism-equivalent up
%   to $\sim$} if there exists an isomorphism $\pi : \str{\tau, D} \rightarrow
%   \str{\tau, D'}$ such that $f (\vec{x}_R) \sim g (\pi \vec{x}_R)$ for all
%   $\vec{x}_R \in \ind(\tau, D)$ .
% \end{definition}

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. We recursively define the equivalence relation
  \emph{syntactic-equivalence}, which we denote using the symbol `$\equiv$', on
  $G$ as follows. Suppose $g$ and $h$ are gates in $C$ such that $\Sigma (g) =
  \Sigma(h)$ and either both $g$ and $h$ are output gates and $\Omega^{-1}(g) =
  \Omega^{-1}(h)$ or neither are output gates. Suppose $g$ and $h$ are input
  gates, then $g \equiv h$ if, and only if, both $g$ and $h$ are constant gates
  or both are relational gates and $\Lambda(g) = \Lambda (h)$. Suppose $g$ and
  $h$ are internal gates and and suppose we have defined the
  syntactic-equivalence relation for all gates of depth less than the depth of
  either $g$ or $h$. Then $g \equiv h$ if, and only if, $L(g) /_\equiv$ and
  $L(h) /_\equiv$ are isomorphism equivalent.
\end{definition}

@@

For a circuit $C$ of order $n$ and a gate $g$ we map think of $g$ as computing
the function that maps an input structure $\mathcal{A}$ and a bijection $\gamma$
from the universe of $\mathcal{A}$ to $[n]$ to the evaluation $C[\gamma
\mathcal{A}(g)$. While we would like to be able to identify gates that compute
the same function in this sense, it is not hard to show that deciding this
equivalence relation for a given circuit is $\NP$-hard. We now show that if two
gates are syntactically-equivalent then the functions computed at these two
gates must be equal. We show later that the syntactic-equivalence relation is
polynomial-time decidable for the class of circuits of interest to us in this
paper.

\begin{lem}
  Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\BB,
  \rho)$-circuit of order $n$. Let $\mathcal{A}$ be a $\rho$-structure of size
  $n$ and let $\gamma$ be a bijection from the universe of $\mathcal{A}$ to
  $[n]$. For all $g, g' \in G$ if $g \equiv g'$ then $C[\gamma \mathcal{A}](g) =
  C[\gamma \mathcal{A}](g')$.
  \label{lem:syntactic-equivalence-equal-function}
\end{lem}
\begin{proof}
  We prove the result by induction on depth. Suppose $g$ and $g'$ have depth $0$
  and $g \equiv g'$. Then they are both input gates and so $g = g'$. The result
  for depth $0$ then follows trivially. Suppose $g$ and $g'$ are internal gates,
  and suppose for all $h, h' \in G$ of depth less than $g$ or $g'$ we have that
  if $h \equiv h'$ then $C[\gamma \mathcal{A}](h) = C[\gamma \mathcal{A}](h')$.
  Suppose $g \equiv g'$. Then exists $\lambda \in \aut(g)$ such that $L(g) (x)
  \equiv L(g') (\lambda x)$ for all $x \in \ind(g)$. It follows from the
  inductive hypothesis that $L^{\gamma \mathcal{A}}(g)(x) = C[\gamma
  \mathcal{A}](L(g)(x)) = C[\gamma \mathcal{A}](L(g')(\lambda x)) = (L^{\gamma
    \mathcal{A}}(g) \lambda) (x)$ for all $x \in \ind(g)$. Since $\Sigma(g)$
  (and so $\Sigma(g')$) is a structured function, it follows that $C[\gamma
  \mathcal{A}](g) = \Sigma(g)(L^{\gamma \mathcal{A}} (g)) = \Sigma(g')
  (L^{\gamma \mathcal{A}})(g') \lambda) = \Sigma(g') (L^{\gamma
    \mathcal{A}}(g')) = C[\gamma \mathcal{A}](g')$. The result follows.
\end{proof}

The syntactic-equivalence relation identifies gates that have `equivalent'
circuits underneath them. A similar intuition is captured by identifying gates
that are mapped to one another by automorphisms of the circuit that extend the
trivial permutation. We now show that if two automorphisms extend the same
permutation then the two images of each gate must be syntactically-equivalent.

% Stated slightly differently, we identify the gates $h$ and $h'$ in a circuit
% $C$ of order $n$ if, and only if, there exists $\sigma \in \sym_n$, $\pi, \pi'
% \in \aut(C)$ both extending $\sigma$, and $g$ a gate in $C$, such that $h =
% \pi (g)$ and $h' = \pi' (g)$. We now show this relation is a refinement of
% syntactic-equivalence.

% \begin{remark}
%   It's interesting to note that if $C$ has unique extensions then this
%   relation collapses to equality, syntactic-equivalence does not. Note to
%   self: should I mention this? Also, the following result is equivalent to
%   checking for just $\pi $
% \end{remark}

@@

  We now prove the result by induction on depth. Suppose $g$ is a gate of depth
  $0$. Then $g$ is either a relational or constant gate. In either case $\pi (g)
  = \pi' (g)$, and so $\pi(g)$ and $\pi'(g)$ are syntactically-equivalent.

  Suppose $g$ is an internal gate and suppose that for every gate $h$ of depth
  less than $g$, $\pi (h)$ is syntactically-equivalent to $\pi'(h)$. We have
  that there exists $\lambda, \lambda' \in \aut(g)$ such that $\pi
  L(g)(\vec{a}_R) = L(\pi g) (\lambda \vec{a}_R)$ and $\pi' L(g)(\vec{a}_R) =
  L(\pi' g)(\lambda' \vec{a}_R)$, for all $\vec{a}_R \in ind(g)$. Then, from the
  inductive hypothesis, we have that $\pi L(g)(\vec{a}_R) \equiv \pi
  'L(g)(\vec{a}_R)$ and so $L(\pi g)(\lambda \vec{a}_R) \equiv L(\pi' g)
  (\lambda' \vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$. Thus we have that
  $L(\pi g) (\vec{a}_R) = L (\pi g) (\lambda \lambda^{-1}\vec{a}_R) \equiv
  L(\pi' g) (\lambda' \lambda^{-1}\vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$,
  and so $L(\pi g)$ is isomorphism equivalent to $L(\pi'g)$ up to
  syntactic-equivalence. We thus have that $\pi (g) $ and $\pi' (g)$ are
  syntactically-equivalent, and the result follows.

@@

It follows from Lemma~\ref{lem:permutation-extending-syntactic-equivalence} that
the syntactic-equivalence relation of a circuit $C$ constrains the automorphism
group of $C$, and so the orbits and stabiliser groups of the gates in $C$. We
say that a circuit $C$ is \emph{syntactically simple} if $C$ has trivial
syntactic-equivalence classes, i.e. for all gates $g$ and $h$ in $C$ if $g
\equiv h$ then $g = h$. An immediate corollary of
Lemma~\ref{lem:permutation-extending-syntactic-equivalence} is that if a circuit
is syntactically simple then it has unique extensions. We now define a few
properties of circuits in terms of the structure of their syntactic-equivalence
classes.

\begin{definition}
  Let $C$ be a circuit and $g$ be a gate in $C$. We say $g$ has \emph{injective
    labels} if $L(g)$ is an injection. We say $g$ has \emph{unique children} if
  no two gates in $H_g$ are syntactically-equivalent. We say $g$ has
  \emph{unique labels} if $g$ has injective labels and unique children. We say
  $C$ has \emph{injective labels} (resp. \emph{unique labels}) if every gate in
  $C$ has injective labels (resp. unique labels). We say $C$ is
  \emph{transparent} if every gate $g$ in $C$ such that $\Sigma(g)$ is not a
  symmetric function has unique labels.
\end{definition}

It is worth noting that if a circuit has injective labels and is syntactically
simple then then it has unique labels. The converse is, in general, false.

@@

We note that the construction of $C'$ from $C$ in
Lemma~\ref{lem:redundant-gates} can be implemented so as to run in time
polynomial in the size of $C$. As such, for the remainder of this paper we
assume, unless stated otherwise, every circuit has no redundant gates. With this
assumption in mind we now show that every circuit with unique children (and so
unique labels) has unique extensions.

@@

\begin{prop}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. If $C$ has unique children then $C$ has unique
  extensions.
  \label{prop:unique-children-unique-extensions}
\end{prop}
\begin{proof}
  Suppose $C$ has unique children. Let $\pi_{\sigma}, \pi_{\sigma}' \in \aut(C)$
  be automorphisms extending $\sigma \in \sym_n$. We now prove that $\pi_\sigma
  = \pi_{\sigma}'$. Let $\pi := \pi_{\sigma}'\pi^{-1}_\sigma$. We have that
  $\pi$ is an automorphism of the circuit and that $\pi$ extends
  $\sigma^{-1}\sigma = e$, the identity permutation. It follows that $\pi$ fixes
  all input gates and, since for any output gate $g$, $\pi (g) = \pi
  (\Omega(\vec{a})) = \Omega(e \vec{a}) = \Omega(\vec{a}) = g$ for all $\vec{a}
  \in [n]^q$, $\pi$ fixes all output gates.

  We now prove the result by structural induction on the circuit starting from
  the output gates. Let $g$ be an output gate then $\pi (g) = g$ from the above
  argument. Let $g$ be an internal, non-output gate. We now show that if $\pi
  (g) = g$ then for all $h \in H_g$ we have $\pi (h) = h$. Suppose $\pi (g) = g$
  and let $h \in H_g$. Since $\pi (g) = g$ we have that $\pi H_{g} = H_g$, and
  so $\pi (h) \in H_g$. We have from
  Lemma~\ref{lem:permutation-extending-syntactic-equivalence} that $h$ is
  syntactically-equivalent to $\pi (h)$ and, since $g$ has unique children, $h$
  is the only gate in $H_g$ syntactically-equivalent to $h$. It follows that
  $\pi (h) = h$.

@@

It follows from Proposition~\ref{prop:unique-children-unique-extensions} that if
$C$ is of order $n$ then for each $\sigma \in \sym_n$ there exists at most one
$\pi_\sigma \in \aut(C)$ extending $\sigma$. However, if $g$ is a gate in $C$
and $\pi_{\sigma} g = g'$ there may no `canonical' $\lambda \in \aut(g)$
witnessing the isomorphism-equivalence, i.e.\ such that $\pi L(g) = L(g')
\lambda$. It is this lack of a `canonical' choice that allows us to prove in
Section~\ref{sec:transparency-and-unique-extensions} that a great many important
circuit properties (e.g. the orbit of a gate, symmetry of a circuit, etc.) for
circuits with unique gates are at least as hard as the graph-isomorphism problem
to decide.

In contrast, it is easy to see that if $C$ has unique labels, and $g$ and $g'$
are gates in the circuit, then $L(g)$ and $L(g')$ are isomorphism-equivalent if,
and only if, $\lambda := L(g')^{-1}L(g)$ is in $\aut(g)$. We show in
Section~\ref{sec:transparency-and-unique-extensions} all of the circuit
properties of interest in this paper are polynomial-time decidable for circuits
with unique labels. Moreover, we show in
Proposition~\ref{prop:transparent-unique} that a transparent circuit can be
converted into an equivalent circuit with unique labels in time polynomial in
the size of the circuit. Since we are working with $P$-uniform families of
transparent circuits, this fact allows us to assume, without a loss of
generality, that transparent circuits have unique labels (and, from
\ref{prop:unique-children-unique-extensions} so unique extensions).
